---
title: "Reshape-11-28"
author: "Shana Arius"
date: "2023-11-28"
output: html_document
---

Let's put this up to github!

```{r}
library(tidyverse)

```


```{r}
gdp <- read_delim("oecd-gdp.csv.bz2")
gdp %>% 
head()
```

long form gdp of aus and chl (using pivot_longer to make it in long form)

```{r}
gdpLong <- 
  gdp %>% 
  pivot_longer(!c(country, indicator))
gdpLong %>% 
  head(10)
```

change the above "name" and "value" columns to "year" and "gdp"
notice that the year column are characters, which could be a potential problem
```{r}
gdpLong <- 
  gdp %>% 
  pivot_longer(!c(country, indicator),
               names_to = "year",
               values_to = "gpd")
gdpLong %>% 
  head(10)
```

to change the year column to numerical values, use as.numeric
```{r}
gdpLong <- 
  gdp %>% 
  pivot_longer(!c(country, indicator),
               names_to = "year",
               values_to = "gpd") %>% 
  mutate(year = as.numeric(year))
gdpLong %>% 
  head(10)

```



make a data frame of a wide form topic
```{r}
p <- data.frame(pregnant = c("Yes", "No"),
            	male = c(NA, 25),
            	female = c(11, 18))
p

```

now transform it to long form 
```{r}
pl <- p %>%
  pivot_longer(c(male, female),
           	names_to = "sex",
           	values_to = "count")
pl


```


now back into wide form bc it's better for certain tasks (but not for things like ggplot); this converts the long form to wide form
```{r}
pl %>% 
  pivot_wider(id_cols = pregnant,
              names_from = sex, 
              values_from = count)
```

